---
title: "R Notebook"
output: html_notebook
---

# Loading and Exploring Data
```{r results = "hide"}

library(tidyverse)
library(corrplot)
library(readr)
library(ggthemes)
library(ggrepel)

train_set <- read_csv("./data/train.csv")

df <- train_set

str(df)

```

# Explore some of the most imporant variables
## Response Variable: SalePrice
``` {r}

df %>% ggplot(aes(x=Id, y=SalePrice)) +
  geom_point() +
  theme_economist_white()

df %>% ggplot(aes(x=SalePrice)) +
  geom_histogram(binwidth = 10000, col="grey") +
  theme_economist_white()

df %>% ggplot(aes(sample=SalePrice)) +
  geom_qq() +
  geom_qq_line() +
  theme_economist_white()

summary(df$SalePrice)

```
## Numerical Predictors
```{r}

numPreds <- df %>% 
  select_if(is.numeric)

cat("Numeric predictors:", length(numPreds))

```
### Correlations with SalePrice
```{r}

cor_numPreds <- numPreds %>% 
  cor(use="pairwise.complete.obs")

# sort the matrix by cor with SalePrice
cor_sorted <- as.matrix(sort(cor_numPreds[,'SalePrice'], decreasing = TRUE))

# select those with > .5 cor with SalePrice
cor_high <- names(which(apply(cor_sorted, 1, function(x) abs(x)>0.5)))
cor_numPreds <- cor_numPreds[cor_high, cor_high]

corrplot.mixed(cor_numPreds, tl.col="black", tl.pos = "lt")

```
In the remainder of this section, we visualize the relation between SalePrice and the two predictors with the highest correlation with SalePrice: Overall Quality and the ‘Above Grade’ Living Area.

Multicollinearity is an issue. For example, the correlation between GarageCars and GarageArea is very high (0.88), and both have similar correlations with SalePrice (0.64, 0.62). The other 6 six variables with a correlation higher than 0.5 with SalePrice are: 

- TotalBsmtSF: Total square feet of basement area 
- 1stFlrSF: First Floor square feet 
- FullBath: Full bathrooms above grade 
- TotRmsAbvGrd: Total rooms above grade (does not include bathrooms) 
- YearBuilt: Original construction date 
- YearRemodAdd: Remodel date (same as construction date if no remodeling or additions)

### Overall Quality
Overall Quality has the highest correlation with SalePrice among the numeric variables (0.79). It rates the overall material and finish of the house on a scale from 1 (very poor) to 10 (very excellent).
```{r}

df %>% ggplot() +
  geom_boxplot(aes(x=factor(OverallQual), y=SalePrice)) + 
  labs(x='Overall Quality') +
  theme_economist_white()

```
<br>

### Above Grade (Ground) Living Area (square feet)
```{r}

df %>% ggplot(aes(x=GrLivArea, y=SalePrice)) +
  geom_point(color = "blue") + 
  geom_text_repel(aes(label = ifelse(GrLivArea > 4500, Id, "")), position = "nudge") +
  theme_economist_white()

```
The two houses with large living areas and low SalePrices seem to be outliers (houses 524 and 1299, see labels in graph). I will not take them out yet, as removing outliers can be dangerous. For instance, a low score on the Overall Quality could explain a low price. However, as you can see below, these two houses score maximum points on Overall Quality. Therefore, I will keep houses 1299 and 524 in mind as prime candidates to remove as outliers.

```{r}

df[c(524, 1299), c("Id", "SalePrice", "GrLivArea", "OverallQual")]

```

## Missing data, label encoding, and factorizing variables
### Completeness
First of all, I would like to see which variables contain missing values.
```{r}

NAcol <- which(colSums(is.na(df)) > 0)
sort(colSums(sapply(df[NAcol], is.na)), decreasing = TRUE)
cat("There are", length(NAcol), "predictors with NA.")

```

### Imputing missing data
```{r}

df$PoolQC <- df$PoolQC %>% replace_na("None")

df$MiscFeature <- df$MiscFeature %>% replace_na("None")

df$Alley <- df$Alley %>% replace_na("None")

df$Fence <- df$Fence %>% replace_na("None")

df$FireplaceQu <- df$FireplaceQu %>% replace_na("None")

#
# replace missing LotFrontages with median of the neighborhood
#
for (i in 1:nrow(df)) {
        if(is.na(df$LotFrontage[i])) {
               df$LotFrontage[i] <- as.integer(median(df$LotFrontage[df$Neighborhood == df$Neighborhood[i]], na.rm=TRUE)) 
        }
}


```



```{r}

library(factoextra)
library(FactoMineR)

df <- readRDS("./rda/df.rds") %>% select(-SalePrice, -trans_SalePrice, -Id) %>% select_if(is.numeric)


df.pca <- PCA(df)
var <- get_pca_var(df.pca)
head(var$contrib, 6)
fviz_contrib(df.pca, choice = "var", axes = 1:2, top = 10)

q <- c("2ndFlrSF", "TotRmsAbvGrd", "GrLivArea", "OverallQual", "GarageArea", "TotalBsmtSF")
df <- readRDS("./rda/df.rds") %>% select(q, SalePrice, trans_SalePrice, Id)
saveRDS(df, file="./rda/mjb1.rds")



```

```{r}
library(caret)

df <- readRDS("./rda/mjb1.rds")

set.seed(1)
test_index <- createDataPartition(df$trans_SalePrice, times = 1, p = 0.2, list = FALSE)
train_set <- df %>% slice(-test_index)
test_set <- df %>% slice(test_index)


```

```{r}

library(randomForest)
library(Rborist)

fit <- train(trans_SalePrice ~ GrLivArea + TotRmsAbvGrd + `2ndFlrSF` + OverallQual, 
             data=train_set, method="Rborist")

summary(fit)

y_hat <- data.frame(Id=test_set$Id, SalePrice_pred = exp(predict(fit, test_set)), SalePrice = test_set$SalePrice)

RMSE(log(y_hat$SalePrice_pred), log(test_set$SalePrice))


y_hat %>% ggplot() +
  geom_point(aes(x = Id, y = SalePrice)) +
  geom_point(aes(x = Id, y = SalePrice_pred), col="red") +
  geom_linerange(aes(x = Id, ymax = SalePrice_pred, ymin = SalePrice)) 




```

